---
title: "Modified_graph_Mat"
output: html_document
date: "2024-02-27"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
source(file = "~/work/Stat_app/DATA/exporting_data_from_link.R")
```

```{r}
#Packages :
install.packages("dplyr")
install.packages("ggplot2")
install.packages("labelled")
library("dplyr")
library("ggplot2")
library("labelled")

library("magrittr")
```


```{r}
# An histogram about the number of hours actually worked in main job
hist(eulfs_small$hwactual, col = "skyblue",
     main = "Distribution of the numbers of hours",
     xlab = "Number of hours",
     ylab = "Number of people")

```
Figure 3. \textbf{The vast majority of Europeans work between 35 and 40 hours a week but those who work the least represent a strong minority.} Histogram of the number of hours actually worked, all years combined. 31 countries from the 2014 release of the EU Labor Force Survey, own analyses.

The peak of the distribution is clearly between 35 and 40 hours with more than 40000 people.
But the second most important interval is between 0 and 5 hours.

--------------------

```{r}
# An histogram about the number of hours that the person would like to work in total in a week
hist(eulfs_small$hwwish, col = "skyblue",
     main = "Distribution of the numbers of hours people wish",
     xlab = "Number of hours",
     ylab = "Number of people")
```
Figure 4. \textbf{The vast majority of Europeans want to work between 35 and 40 hours a week.} Histogram of the number of hours people wish to work, all years combined. 31 countries from the 2014 release of the EU Labor Force Survey, own analyses.

By contrast with the number of hours actually worked, there is almost no one who want to work less than 5 hours a week. Europeans tend to be unanimous on this issue.

---------------

```{r}
# Boxplot to compare men and women concerning the number of hours actually worked in main job
boxplot(eulfs_small$hwactual ~ eulfs_small$sex,
        xlab = " ",
        ylab = "Number of hours",
        names = c("Men", "Women"),
        col = c("blue", "red"),
        main = "Actual weekly working hours: men and women"
)

```
Figure 4. \textbf{Although the median number of working hours is almost the same for each sex, the variability is much higher for women.} Boxplot of the number of hours actually worked depending on sex, all years combined. 31 countries from the 2014 release of the EU Labor Force Survey, own analyses.

The variability is much higher for female workers: their first quartile that is close from the minimum of men.

---------------

We may analyse data for a specific country, for instance Austria.

```{r fig.cap= ""}
# Starting by creating a smaller base to work with
eulfs_AT <- eulfs_small %>% filter(country == 1) %>% select(hwactual, refyear)
# Indicating what statistics we want to appear on the plot
frame <- eulfs_AT %>% group_by(refyear, .drop = NA) %>% summarise(Min = min(hwactual, na.rm = TRUE), Max = max(hwactual, na.rm = TRUE), Median = median(hwactual, na.rm = TRUE), quant1 = quantile(hwactual, probs = 0.25, na.rm = TRUE), quant2 = quantile(hwactual, probs = 0.75, na.rm = TRUE))
# Creating the graph
p <- ggplot(frame, aes(x = refyear, na.rm = TRUE)) + geom_line(aes(y=Min)) + geom_line(aes(y=Max), color = "green") + geom_line(aes(y=Median), color = "darkred") + geom_line(aes(y=quant1), color = "blue") + geom_line(aes(y=quant2), color = "red") + labs(title = "Actual weekly working hours in 1st job in Austria", x=" ", y="Number of weekly working hours")
# Plotting the graph
p
```
Figure X. \textbf{Those who work the least tend to have less and less working hours in recent years in Austria.} A few order statistics concerning Austria: the minimum, the three quartiles and the maximum of the number of hours actually worked in main job. From the 2014 release of the EU Labor Force Survey, own analyses.

We notice that all quartiles remain somewhat stable, except the first quartile.
Indeed, the quarter of the population who work the least has less and less working hours.
The number of hours dropped below 20 hours after 2010.

