---
title: "Oaxaca-dec1omposition : explaining the differences in hours usually                   worked per week over the years"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r cars}
library("dplyr")
library("ggplot2")
library("labelled")
library("haven")
library("tidyr")
library("plm")
library("gtsummary")
library(oaxaca)
library(openxlsx)
library(tinytex)

```



```{r pressure, echo=FALSE}
source(file = "~/Stat_app/DATA/exporting_data_from_link.R")

```

```{r pressure, echo=FALSE}
path_occ <- "~/Stat_app/Occupation/occupation_status.xlsx"

# Lire le fichier Excel
occu_code <- read.xlsx(path_occ)
```

```{r pressure, echo=FALSE}

#Nettoyage de la base de données :

dec1 <- eulfs_small
#1. Enlever les valeurs NA de la variable expliquée : 

dec1$hwusual<-na_if(dec1$hwusual,99)
dec1 <-dec1%>% filter(!is.na(hwusual))

#2. Variable education : 

dec1$hatlev1d<-na_if(dec1$hatlev1d,99)
dec1 <-dec1%>% filter(!is.na(hatlev1d))

#On fait de la variable hatlev1d 3 variables binaires :

dec1$high_educ <- ifelse(dec1$hatlev1d==3,1,0)
dec1$medium_educ <- ifelse(dec1$hatlev1d==2,1,0)
dec1$low_educ <- ifelse(dec1$hatlev1d==1,1,0)

#3. Occupation status : 
#Création d'un nouveau df pour récupérer les variables occupation 

occupation<-subset(eulfs_small, select=c("year","isco1d","is881d"))
occupation$isco1d<-na_if(occupation$isco1d,99)
occupation$is881d<-na_if(occupation$is881d,99)
occupation$isco1d<-ifelse(occupation$year > 2010 & is.na(occupation$isco1d),18,
                          occupation$isco1d)
occupation$is881d<-ifelse(occupation$year < 2011 & is.na(occupation$is881d),18,
                          occupation$is881d)

occupation$occupation_continuous <- ifelse(is.na(occupation$isco1d),occupation$is881d,occupation$isco1d)
df<-eulfs_small
df$occupation_continous<-occupation$occupation_continuous

df$hatlev1d<-na_if(df$hatlev1d,99)
df<-df %>% filter(!is.na(hatlev1d))
df$hwusual<-na_if(df$hwusual,99)
df<-df %>% filter(!is.na(hwusual))

dec1$occupation <- df$occupation_continous
dec1<- dec1 %>% filter(occupation!= 18)

#On crée une variable binaire pour chaque statut

dec1$armed <- ifelse(dec1$occupation==0,1,0)
dec1$manager <- ifelse(dec1$occupation==100,1,0)
dec1$pro <- ifelse(dec1$occupation==200,1,0)
dec1$tech <- ifelse(dec1$occupation==300,1,0)
dec1$cler <- ifelse(dec1$occupation==400,1,0)
dec1$service <- ifelse(dec1$occupation==500,1,0)
dec1$agri <- ifelse(dec1$occupation==600,1,0)
dec1$trade <- ifelse(dec1$occupation==700,1,0)
dec1$operator <- ifelse(dec1$occupation==800,1,0)
dec1$elementary <- ifelse(dec1$occupation==900,1,0)

#4. Gender

dec1$women <-if_else(dec1$sex==2,1,0)

```

```{r}
#Construire nos différentes bases pour nos différentes décompositions : 

oaxaca_year1 <- dec1 %>% subset(select=c("year","hwusual","low_educ","high_educ"
,"medium_educ","armed", "manager","pro","tech","cler","service","agri","operator"
,"elementary","occupation","women","trade"))
oaxaca_1 <- oaxaca_year1%>%filter(year==1998|year==1999|year==2000|year==2011|year==2012|year==2013)
oaxaca_1$y13 <- if_else(oaxaca_1$year==2011|oaxaca_1$year==2012|oaxaca_1$year==2013,1,0)

```

```{r pressure, echo=TRUE}

#First dec1omposition : 1998-2013
diff_2 <- oaxaca(formula = hwusual ~ women + high_educ + low_educ + manager 
                      + pro + tech + cler + service + + agri + trade + operator 
                      + elementary |y13|manager + pro +
                        tech + cler + service + + agri + trade + operator +
                        elementary, data=oaxaca_1, R=1000)
```

```{r pressure, echo=FALSE}
#Observations 
diff_2$n



#Differences :
diff_2$y


```
In our analyzed sample we have 17476 individuals working in 1998, 1999 and 2000
and 20299 working in 2011,2012 and 2013.
The y component of the resulting oaxaca-class object indicates that the mean of
hours worked per week is 37.5 hours for group A (1998,1999 and 2000) and 
36.7 hours for group B (2011,2012 and 2013) leaving 
the difference of approximately 0.8 hours to be explained by the Blinder-Oaxaca
dec1omposition.

**Threefold dec1omposition :**

```{r pressure, echo=FALSE}

diff_2$threefold$overall

```

The results suggest that 0,1 hours of the 0,8 hours difference can be 
attributed to group-differences in endowments (gender,education), 0,7 to 
differences in coefficients, and 0,1 to the interaction between the two.



  
```{r pressure, echo=FALSE}

plot(diff_2, dec1omposition = "threefold",
     components = c("endowments","coefficients"), 
     unexplained.split = TRUE,
     variables=c("women","high_educ","low_educ", "manager","tech","pro", "cler", 
                 "service", "agri","trade", "operator","elementary"),
    variable.labels=c("women"="Women","high_educ"="High level of education", 
                      "low_educ"="Low level of education", "manager"="Managers",
                      "tech"="Technicians and associate professionals",
                      "cler"="Clerical support workers",
                      "pro"="Professionals",  "cler"="Clerical support workers", 
                      "service"=" Service and sales workers", 
                      "agri"=" Skilled agricultural, forestry and fishery workers", 
                      "trade"="Craft and related trades workers",
                      "operator"="Plant and machine operators and assemblers",     
                      "elementary"="Elementary occupations"))
```
*Figure 1.* The endowments and coefficients components of a threefold 
Blinder-Oaxaca dec1omposition of the 1998 vs. 2013 working hours gap.

*Endowments*

Figure 1 shows the estimation results for each variable, along with error bars 
that indicate 95% confidence intervals. 
In the endowments component, most variables appear to have a
statistically insignificant (or only marginally significant) influence, 
with the exceptions of
the variables Women, Low Level of education, High Level of education, 
Professionals, Service and Salesworkers,Craft and related trades workers, 
Plant and machine operators 
and assemblers and Managers.
Therefore, the variables explaining the mean difference are : Women, 
Professionals,Craft and related trades workers, Plant and machine operators 
and assemblers and Managers.

```{r pressure, echo=FALSE}


print("women")
summary(diff_2$reg$reg.pooled.2)$coefficients["women",]
print("manager")
summary(diff_2$reg$reg.pooled.2)$coefficients["manager",]
print("pro")
summary(diff_2$reg$reg.pooled.2)$coefficients["pro",]
print("trade")
summary(diff_2$reg$reg.pooled.2)$coefficients["trade",]
print("service")
summary(diff_2$reg$reg.pooled.2)$coefficients["service",]
print("operator")
summary(diff_2$reg$reg.pooled.2)$coefficients["operator",]


```
The only coefficients that have a significant p-value are "women" and "manager".

```{r pressure, echo=TRUE}
diff_2$x$x.mean.diff["women"]
```
There are more women working in the years 2011,2012 and 2013 than in the years
1998,1999 and 2000. The pooled regressions shows that being a woman dec1reases 
the mean of working hours by 4,5 hours. 

```{r pressure, echo=TRUE}

diff_2$x$x.mean.diff["manager"]
```
A larger proportion of individuals working in 1998/1999/2000 work in the 
Managers occupational status. However the pooled regressions shows us that this 
occupational status works 4,6 hours more than the average. 


*Coefficients :*

```{r}

diff_2$beta$beta.diff["manager"]
   
```
The difference in coefficients of manager drives a large 
proportion  of the gap between 1998/1999/2000 vs. 2011/2012/2013. We understand 
that the return in working hours of working in the Managers staus,is greater in 
the years 1998/1999/2000 and 2011/2012/2013.

*Twofold dec1omposition*

```{r pressure, echo=TRUE}
diff_2$twofold$overall
```
We focus our discussion on the Neumark dec1omposition, which
uses pooled regression coefficients, denoted by -1 in the weights column, 
(from a regression that does not include the group indicator variable of 2013) 
as the reference 
coefficient set, . The results of the overall twofold dec1omposition indicate 
that the working hours gap between 1998/1999/2000 and 2011/2012/2013 workers can 
be dec1omposed into
0,4 hours that can be explained by group differences in the explanatory 
variables and 2,4 hours that is unexplained.

```{r pressure, echo=TRUE}

plot(diff_2, decomposition = "twofold",
     components = c("explained","unexplained"), 
     group.weight = -1,
     variables=c("women","high_educ","low_educ", "manager","tech","pro", "cler", 
                 "service", "agri","trade", "operator","elementary"),
    variable.labels=c("women"="Women","high_educ"="High level of education", 
                      "low_educ"="Low level of education", "manager"="Managers",
                      "tech"="Technicians and associate professionals",
                      "cler"="Clerical support workers",
                      "pro"="Professionals",  "cler"="Clerical support workers", 
                      "service"=" Service and sales workers", 
                      "agri"=" Skilled agricultural, forestry and fishery workers", 
                      "trade"="Craft and related trades workers",
                      "operator"="Plant and machine operators and assemblers",     
                      "elementary"="Elementary occupations"))
```
*Figure 2* The explained and unexplained components of a twofold Blinder-Oaxaca
dec1omposition of the 1998/1999/2000 vs. 2011/2012/2013 working hours gap.
The results are consistent with the threefold composition. 
```{r pressure, echo=TRUE}
plot(diff_2, decomposition = "twofold", group.weight = -1,
     unexplained.split = TRUE,
     components="explained",
     variables=c("women","high_educ","low_educ", "manager","tech","pro", "cler", 
                 "service", "agri","trade", "operator","elementary"),
    variable.labels=c("women"="Women","high_educ"="High level of education", 
                      "low_educ"="Low level of education", "manager"="Managers",
                      "tech"="Technicians and associate professionals",
                      "cler"="Clerical support workers",
                      "pro"="Professionals",  "cler"="Clerical support workers", 
                      "service"=" Service and sales workers", 
                      "agri"=" Skilled agricultural, forestry and fishery workers", 
                      "trade"="Craft and related trades workers",
                      "operator"="Plant and machine operators and assemblers",     
                      "elementary"="Elementary occupations"))
```

```{r pressure, echo=TRUE}
plot(diff_2, decomposition = "twofold", group.weight = -1,
     components="unexplained",
     variables=c("women","high_educ","low_educ", "manager","tech","pro", "cler", 
                 "service", "agri","trade", "operator","elementary"),
    variable.labels=c("women"="Women","high_educ"="High level of education", 
                      "low_educ"="Low level of education", "manager"="Managers",
                      "tech"="Technicians and associate professionals",
                      "cler"="Clerical support workers",
                      "pro"="Professionals",  "cler"="Clerical support workers", 
                      "service"=" Service and sales workers", 
                      "agri"=" Skilled agricultural, forestry and fishery workers", 
                      "trade"="Craft and related trades workers",
                      "operator"="Plant and machine operators and assemblers",     
                      "elementary"="Elementary occupations"))
```


```{r pressure, echo=TRUE}



plot(diff_2, decomposition = "twofold", group.weight = -1,
     unexplained.split = TRUE,
     components=c("unexplained A","unexplained B"),
     component.labels = c("unexplained A"="Unexplained A",
                          "unexplained B"="Unexplained B"),
     variables=c("high_educ", "manager"),
    variable.labels=c("high_educ"="High level of education", "manager"="Managers"),
    component.left=TRUE)

```
*Figure 4* The unexplained portion’s discrimination sub-components in a 
twofold Blinder-
Oaxaca dec1omposition of the 1998 vs. 2013 working hours gap. 


```{r}
variables <- c("low_educ", "manager")
columns <- c("group.weight", "coef(unexplained A)", "coef(unexplained B)")
diff_2$twofold$variables[[5]][variables, columns]

```



*Conclusion*
The results suggests that much of the gap between 1998/1999/2000 and 
2011/2012/2013 can be explained by the following facts : 
-There are less workers working in the occupational status of "Managers" and 
when people working in this professionals statu tends to work more hours than 
the average.
-There are more workers working in the occupational status of "Professionals", 
"Service and sales workers" and more women working in the last three years than 
in the first three years. Those workers appears to 
have a mean of working hours lower than the overall mean.
-Working in the professional status managers increases working hours in a bigger 
way in 1998/1999/2000 than it does in 2011/2012/2013. 






