---
title: "Austria"
output:
  pdf_document: default
  html_document: default
date: "2024-05-10"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
source(file = "~/work/Stat_app/DATA/exporting_data_from_link.R")
```

```{r}
source(file="~/work/Stat_app/Countries/countries_code.R")
```

```{r}
#Packages :
install.packages("dplyr")
install.packages("ggplot2")
install.packages("labelled")
library("dplyr")
library("ggplot2")
library("labelled")

# Add this package for %>%
library("magrittr")
```

```{r}
install.packages("reshape2")
```



We may analyse data for a specific country, for instance Austria.

```{r fig.cap= ""}
# Starting by creating a smaller base to work with
eulfs_AT <- eulfs_small %>% filter(country == 1) %>% select(hwusual, refyear)
# Indicating what statistics we want to appear on the plot
frame <- eulfs_AT %>% group_by(refyear, .drop = NA) %>% summarise(Min = min(hwusual, na.rm = TRUE), Max = max(hwusual, na.rm = TRUE), Median = median(hwusual, na.rm = TRUE), quant1 = quantile(hwusual, probs = 0.25, na.rm = TRUE), quant2 = quantile(hwusual, probs = 0.75, na.rm = TRUE))
# Creating the graph
p <- ggplot(frame, aes(x = refyear, na.rm = TRUE)) + geom_line(aes(y=Min)) + geom_line(aes(y=Max), color = "green") + geom_line(aes(y=Median), color = "darkred") + geom_line(aes(y=quant1), color = "blue") + geom_line(aes(y=quant2), color = "red") + labs(title = "Usual weekly working hours in main job in Austria", x=" ", y="Number of weekly working hours")
# Plotting the graph
p
```
Figure 7. **Those who work the least tend to have less and less working hours in recent years in Austria**. A few order statistics concerning Austria: the minimum, the three quartiles and the maximum of the number of hours usually worked in main job. From the 2014 release of the EU Labor Force Survey, own analyses.

We notice that all quartiles remain somewhat stable, except the first quartile.
Indeed, the quarter of the population who work the least has less and less working hours.
The number of hours dropped below 20 hours after 2010.


