---
title: "Oaxaca-Blinder decomposition"
output: html_document
date: "2024-02-27"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# First getting the eulfs file
source(file = "~/work/Stat_app/DATA/exporting_data_from_link.R")
```

```{r}
# Usual packages
install.packages("dplyr")
install.packages("ggplot2")
install.packages("labelled")
library("dplyr")
library("ggplot2")
library("labelled")

# Specific package
install.packages("oaxaca")
```

```{r}
# To avoid problems with 99
eulfs_small$hwactual <- na_if(eulfs_small$hwactual, 99)
eulfs_small$hwwish <- na_if(eulfs_small$hwwish, 99)
```

```{r}
# Creating a function to summarize what we learn thanks to Oaxaca decomposition
Blinder_Oaxaca <- function(group0, group1, res)
  # group0 and group1: names of the groups involved in the decomposition, res: results obtained thanks to oaxaca::oaxaca
{ 
# Number of observations
print(paste("Among", res$n[[3]], "observations, we have", res$n[[1]], group0, "and", res$n[[2]], group1, "."))
# Looking at the means
print(paste("The mean of the number of hours actually worked is", res$y[[1]], "hours for", group0, "and", res$y[[2]], "for", group1, "leaving a difference of approximately", res$y[[3]], "hours to be explained by the Blinder-Oaxaca decomposition."))
# Exploring the decomposition
print(paste("This decomposition suggests that, of the", res$y[[3]], "hours difference, approximately", res$threefold$overall[[1]], "can be attributed to group differences in endowments (i.e., age, education, living in a city or not),", res$threefold$overall[[3]], "to differences in coefficients, and the remaining", res$threefold$overall[[5]] , "is accounted for by the interaction of the two."))
  }

```

## One basic example: men and women
```{r}
# Slightly modifying our dataset before proceding to the decomposition
eulfs_small_bis <- eulfs_small
eulfs_small_bis$log_sex <- as.logical(eulfs_small_bis$sex - 1)
# To have an indicator variable to distinguish two groups
# FALSE is for male, TRUE for female
```

```{r}
# The decomposition itself
results <- eulfs_small_bis %>% oaxaca::oaxaca(formula = hwactual ~ age + hatlev1d + as.factor(degurba) | log_sex)

# Applying the function to this specific case
Blinder_Oaxaca("men", "women", results)
```

## Visualizing the decomposition

```{r, echo = FALSE}
plot(results, components = c("endowments","coefficients"), variable.labels = c("age" = "Age", "hatlev1d" = "Education", "as.factor(degurba)2" = "Living in towns and suburbs", "as.factor(degurba)3" = "Living in rural areas"))
```
Figure 1: *Though men and women seem to have similar endowments, these are not valued the same as male ones.* The endowments and coefficients components of a threefold Blinder-Oaxaca decomposition of the men vs. women working hours gap.

We may see the importance of each type of coefficients.
Globally, we see that the endowments can only explain a very small part of the gap, most of it remains linked to how much endowments are valued for each sex.
