---
title: "Recentered Influence Functions"
output: pdf_document
date: "2024-03-29"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
install.packages("dplyr")
install.packages("ggplot2")
install.packages("labelled")
install.packages("dineq") # specific package for RIF
library("dplyr")
library("ggplot2")
library("labelled")
```

```{r}
source(file = "~/work/Stat_app/DATA/exporting_data_from_link.R")
```

```{r}
eulfs_small$hwusual <- na_if(eulfs_small$hwusual, 99)
```

Source:
https://search.r-project.org/CRAN/refmans/dineq/html/00Index.html

## Recentered influence function of the 1st quartile (25th quantile)

```{r}
rif_q25 <- dineq::rif(x=eulfs_small$hwusual, weights=NULL, method="quantile", quantile=0.25)
```

rif_q25 is a vector with as many components as observations in eulfs_small
Each component corresponds to a particular individual’s influence on the distributional statistic

## Regression with some RIF

```{r}
# RIF Regression with the same distributional statistic, namely the 25th quantile
rifr_q <- dineq::rifr(hwusual ~ sex, eulfs_small, weights = NULL, method = "quantile", quantile = 0.25, kernel = "gaussian")
```

```{r}
rifr_q[["Coef"]]
```

```{r}
plot(eulfs_small$hwusual, eulfs_small$sex)+
  abline(rifr_q)
# Aucun interet
```

```{r}

```


## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
